\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-09-21 analysis/paper/paper.Rmd
added casebase package to intro, and vignette to discussion

In this article, we present a comprehensive and flexible \proglang{R} package for parametric survival analysis that combines the ideas of Hanley & Miettinen [-@hanley2009fitting] into a simple interface. The purpose of the \pkg{casebase} package is to provide practitioners with an easy-to-use software tool to compute a patient's risk (or cumulative incidence) of an event, conditional on a particular patient's covariate profile. Our package retains the flexibility of case-base sampling and the familiar interface of the \code{glm} function. It also provides tools for variable selection and competing-risk analysis. In addition, we provide extensive visualization tools.

@@

\begin{table}[H]

@@

For our first case study, we return to the ERSPC study and investigate the differences in risk between the control and screening arms. We fit four models that differ in which functional form of time is used: 1) excluded from the linear predictor as seen in \eqref{eq:exp}, 2) linear function as seen in \eqref{eq:gomp}, 3) log function as seen in \eqref{eq:weibull}, and 4) a smooth function using cubic B-splines. The models are fit using \code{fitSmoothHazard} with the familiar formula interface:

@@

@@

In conclusion, we presented the \proglang{R} package \pkg{casebase} which implements case-base sampling for fitting parametric survival models and for estimating smooth cumulative incidence functions using the framework of generalized linear models. We strongly believe that its flexibility and its foundation on the familiar logistic regression model will make it appealing to new and established practitioners.

@@

We would like to thank Yi Yang for helpful discussions on penalized regression models. \mbox{Bhatnagar} gratefully acknowledges funding via a Discovery Grant from the Natural Sciences and Engineering Research Council of Canada (NSERC).

>>>>

In this article, we present the \pkg{casebase} package [@casebase-package] implemented in \proglang{R} [@r-core], for parametric survival analysis that combines the ideas of Hanley & Miettinen [-@hanley2009fitting] into a simple interface. The purpose of the \pkg{casebase} package is to provide practitioners with an easy-to-use software tool to compute a patient's risk (or cumulative incidence) of an event, conditional on a particular patient's covariate profile. Our package retains the flexibility of case-base sampling and the familiar interface of the \code{glm} function. It also provides tools for variable selection and competing-risk analysis. In addition, we provide extensive visualization tools.

@@

\begin{table}[ht]

@@

For our first case study, we return to the ERSPC study and investigate the differences in risk between the control and screening arms.

\subsection{Different functional forms of time}

We fit four models that differ in which functional form of time is used: 1) excluded from the linear predictor as seen in \eqref{eq:exp}, 2) linear function as seen in \eqref{eq:gomp}, 3) log function as seen in \eqref{eq:weibull}, and 4) a smooth function using cubic B-splines. The models are fit using \code{fitSmoothHazard} with the familiar formula interface:

@@

\subsection{Time-dependent hazard ratios}

@@

In conclusion, we presented the \proglang{R} package \pkg{casebase} which implements case-base sampling for fitting parametric survival models and for estimating smooth cumulative incidence functions using the framework of generalized linear models. We strongly believe that its flexibility and its foundation on the familiar logistic regression model will make it appealing to new and established practitioners. The \pkg{casebase} package is freely available from the Comprehensive \proglang{R} Archive Network at https://cran.r-project.org/package=casebase. Interested users can visit http://sahirbhatnagar.com/casebase/ for detailed package documentation and vignettes.

@@

We would like to thank Yi Yang for helpful discussions on penalized regression models and gradient boosting. \mbox{Bhatnagar} gratefully acknowledges funding via a Discovery Grant from the Natural Sciences and Engineering Research Council of Canada (NSERC).

### 2020-09-21 analysis/paper/paper.tex
added casebase package to intro, and vignette to discussion

In this article, we present a comprehensive and flexible \proglang{R}
package for parametric survival analysis that combines the ideas of
Hanley \& Miettinen \citeyearpar{hanley2009fitting} into a simple
interface. The purpose of the \pkg{casebase} package is to provide
practitioners with an easy-to-use software tool to compute a patient's
risk (or cumulative incidence) of an event, conditional on a particular
patient's covariate profile. Our package retains the flexibility of
case-base sampling and the familiar interface of the \code{glm}
function. It also provides tools for variable selection and
competing-risk analysis. In addition, we provide extensive visualization
tools.

@@

In conclusion, we presented the \proglang{R} package \pkg{casebase}
which implements case-base sampling for fitting parametric survival
models and for estimating smooth cumulative incidence functions using
the framework of generalized linear models. We strongly believe that its
flexibility and its foundation on the familiar logistic regression model
will make it appealing to new and established practitioners.

@@

This report was generated on 2020-09-21 12:49:15 using the following
computational environment and dependencies:

>>>>

In this article, we present the \pkg{casebase} package
\citep{casebase-package} implemented in \proglang{R} \citep{r-core}, for
parametric survival analysis that combines the ideas of Hanley \&
Miettinen \citeyearpar{hanley2009fitting} into a simple interface. The
purpose of the \pkg{casebase} package is to provide practitioners with
an easy-to-use software tool to compute a patient's risk (or cumulative
incidence) of an event, conditional on a particular patient's covariate
profile. Our package retains the flexibility of case-base sampling and
the familiar interface of the \code{glm} function. It also provides
tools for variable selection and competing-risk analysis. In addition,
we provide extensive visualization tools.

@@

In conclusion, we presented the \proglang{R} package \pkg{casebase}
which implements case-base sampling for fitting parametric survival
models and for estimating smooth cumulative incidence functions using
the framework of generalized linear models. We strongly believe that its
flexibility and its foundation on the familiar logistic regression model
will make it appealing to new and established practitioners. The
\pkg{casebase} package is freely available from the Comprehensive
\proglang{R} Archive Network at
\url{https://cran.r-project.org/package=casebase}. Interested users can
visit \url{http://sahirbhatnagar.com/casebase/} for detailed package
documentation and vignettes.

@@

This report was generated on 2020-09-21 14:06:21 using the following
computational environment and dependencies:
